Drivers of sex differences in consumption

### 2017-06-28 sex-difference/namcs.Rmd
sex diff

```{r visit_total_odds}
x = svymean(~sex, design=adult_design) %>%
  as_data_frame() %>%
  mutate(sex=c('female', 'male'),
         odds=mean/min(mean))

female_visit_odds = x %>% filter(sex=='female') %$% odds

x %>%
  select(sex, `fraction of visits`=mean, SE, odds) %>%
  kable
```

Overall, more of the visits were made by women. Women make `r round(female_visit_odds - 1, 2)*100`% more visits than men. I'll therefore evaluate consumption by two standards: first, do women consume more antibiotics overall; second, do women consume more antibiotics *per visit*.

@@

abx_total_f = function(abx) {
  svyby(as.formula(paste0('~', abx)), ~sex, adult_design, svytotal) %>%
    rename_(.dots=setNames(list(abx), c('total'))) %>%
    select(sex, total, se) %>%
    mutate(abx=abx)
}

poisson_f = function(abx) {
  svyglm(as.formula(paste0(abx, '~sex')), adult_design) %>%
    tidy %>%
    filter(term != '(Intercept)') %>%
    mutate(abx=abx) %>%
    select(abx, term, estimate, p.value)
}

@@

```{r total_plot}
abx_totals %>%
  mutate_at(vars(total, se), mpy) %>%
  mutate(ci=se*1.96,
         ci_low=total-ci,
         ci_high=total+ci) %>%
  select(abx, sex, total, ci_low, ci_high) %>%
  gather('key', 'value', total:ci_high) %>%
  mutate(key=case_when(.$sex=='male' & .$key=='total' ~ 'x',
                       .$sex=='male' & .$key=='ci_low' ~ 'xmin',
                       .$sex=='male' & .$key=='ci_high' ~ 'xmax',
                       .$sex=='female' & .$key=='total' ~ 'y',
                       .$sex=='female' & .$key=='ci_low' ~ 'ymin',
                       .$sex=='female' & .$key=='ci_high' ~ 'ymax')) %>%
  select(-sex) %>%
  spread(key, value) %>%
  mutate(highlight=abx %in% c('nitro', 'clinda')) %>%
  ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=abx, color=highlight)) +
  geom_abline(col='gray') + geom_abline(slope=female_visit_odds, col='gray') +
  geom_point() + geom_errorbar() + geom_errorbarh() +
  scale_color_manual(values=c('black', 'red')) +
  geom_text() +
  #geom_text(nudge_x=1, nudge_y=0.5) +
  #scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) + xlab('No. rx. for men (millions)') +
  #scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 14)) + ylab('No. rx. for women (millions)') +
  coord_fixed() +
  theme_minimal()
```

@@

abx_diff %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
  mutate(sig=case_when(.$adj.p.value < 0.05 ~ '**',
                       .$p.value < 0.05 ~ '*',
                       TRUE ~ '')) %>%
  mutate_if(is.numeric, scientific_round) %>%
  kable
```

@@

```{r pois_table}
abx_pois %>%
  select(abx, estimate, p.value) %>%
  arrange(p.value) %>%
  mutate(p.value.adj=p.adjust(p.value, method='BH')) %>%
  mutate_at(vars(-abx), scientific_round) %>%
  kable
```

@@

@@

>>>>

```{r visit_total_odds}
x = svymean(~sex, design=adult_design) %>%
  as_data_frame() %>%
  mutate(sex=c('female', 'male'))

odds = function(x) x/(1-x)

p = x %>% filter(sex=='female') %$% mean
se = x %>% filter(sex=='female') %$% SE
female_visit_odds = odds(c(p-se, p, p+se))

x %>%
  select(sex, `fraction of visits`=mean, SE) %>%
  kable
```

Overall, more of the visits were made by women. Women make `r round(female_visit_odds[2], 3)*100`% more visits than men. I'll therefore evaluate consumption by two standards: first, do women consume more antibiotics overall; second, do women consume more antibiotics *per visit*.

@@

abx_total_f = function(abx, design=adult_design) {
  svyby(as.formula(paste0('~', abx)), ~sex, design, svytotal) %>%
    rename_(.dots=setNames(list(abx), c('total'))) %>%
    select(sex, total, se) %>%
    mutate(abx=abx)
}

poisson_f = function(abx, design=adult_design) {
  svyglm(as.formula(paste0(abx, '~sex')), design) %>%
    tidy %>%
    filter(term != '(Intercept)') %>%
    mutate(abx=abx) %>%
    select(abx, term, estimate, p.value)
}

@@

```{r total_plot}
total_plot_f = function(df, highlight=character(0)) {
  df %>%
    mutate_at(vars(total, se), mpy) %>%
    mutate(ci=se*1.96,
           ci_low=total-ci,
           ci_high=total+ci) %>%
    select(abx, sex, total, ci_low, ci_high) %>%
    gather('key', 'value', total:ci_high) %>%
    mutate(key=case_when(.$sex=='male' & .$key=='total' ~ 'x',
                         .$sex=='male' & .$key=='ci_low' ~ 'xmin',
                         .$sex=='male' & .$key=='ci_high' ~ 'xmax',
                         .$sex=='female' & .$key=='total' ~ 'y',
                         .$sex=='female' & .$key=='ci_low' ~ 'ymin',
                         .$sex=='female' & .$key=='ci_high' ~ 'ymax')) %>%
    select(-sex) %>%
    spread(key, value) %>%
    mutate(highlight=abx %in% highlight) %>%
    ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=abx, color=highlight)) +
    geom_abline(col='gray') +
    geom_point() + geom_errorbar() + geom_errorbarh() +
    scale_color_manual(values=c('black', 'red')) +
    geom_text() +
    #geom_text(nudge_x=1, nudge_y=0.5) +
    #scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) +
    #scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 14)) +
    xlab('No. rx. for men (millions)') + ylab('No. rx. for women (millions)') +
    coord_fixed() +
    theme_minimal()
}

total_plot_f(abx_totals, highlight=c('nitro', 'clinda')) +
  geom_abline(slope=female_visit_odds[2], col='gray') +
  geom_abline(slope=female_visit_odds[-2], col='gray', linetype='dotted')
```

@@

sig_f = function(df) {
  stopifnot('adj.p.value' %in% names(df) && 'p.value' %in% names(df))
  case_when(df$adj.p.value < 0.05 ~ '**',
            df$p.value < 0.05 ~ '*',
            TRUE ~ '')
}

abx_diff %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
  mutate(sig=sig_f(.)) %>%
  mutate_if(is.numeric, scientific_round) %>%
  kable
```

@@

```{r pois_table}
abx_pois %>%
  select(abx, estimate, p.value) %>%
  arrange(p.value) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
  mutate(sig=sig_f(.)) %>%
  mutate_if(is.numeric, scientific_round) %>%
  kable
```

@@

## Just among the elderly

```{r elderly}
elderly_design = subset(design, age_group == 'elderly')

elderly_usage = svymean(~sex, design=elderly_design) %>%
  as_data_frame() %>%
  mutate(sex=c('female', 'male'))

elderly_female_odds = elderly_usage %>% filter(sex=='female') %$% mean %>% odds

elderly_totals = abx_totals = lapply(top_abx$abx, function(x) abx_total_f(x, design=elderly_design)) %>% bind_rows()
total_plot_f(elderly_totals) +
  geom_abline(slope=elderly_female_odds, color='gray')
```

Usage among the elderly is more even: women make `r 100*round(elderly_female_odds - 1, digits=3)`% more visits than men. Patterns of consumption appear to be a little difference: more ciprofloxacin, azithro is more concentrated among women, and some drugs (e.g., doxy) appear to approach the line of equal consumption.

@@

Among the elderly, patterns of consumption appear to be somewhat different, and those patterns are not exactly what we see in the adult NAMCS data or even in the Medicare data I have.

### 2017-06-28 sex-difference/sex-difference.Rmd
sex diff

@@

```{r drugs_by_sex}
usage_by_sex = function(drug) {
  bene %>%
    select(-n_claims) %>%
    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id) %>% rename(n_claims=n), by='bene_id') %>%
    replace_na(list(n_claims=0)) %>%
    group_by(sex) %>%
    summarize(mean_n_claims=mean(n_claims)) %>%
    ungroup() %>%
    spread(sex, mean_n_claims) %>%
    mutate(drug=drug)
}

dbs = lapply(unique(pde$antibiotic), usage_by_sex) %>% bind_rows
```

@@

```{r drug_by_sex_plot}
ubs %>%
  mutate(highlight=antibiotic %in% c('azithromycin', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'fosfomycin', 'nitrofurantoin')) %>%
  ggplot(aes(x=male, y=female, label=antibiotic, color=highlight)) +
  scale_color_manual(values=c('black', 'red')) +
  geom_text() +
  geom_point() +
  geom_abline() +
  xlab('claims per 1,000 male beneficiaries') +
  ylab('claims per 1,000 female beneficiaries') +
  theme_minimal() +
  coord_fixed()
```

@@

```{r}
group_by_all_ = function(df, group_name) {
  bind_rows(df, mutate_(df, .dots=setNames(list(~'all'), group_name))) %>%
    group_by_(group_name)
}

epi %>%
  group_by_all_('sex') %>%
  summarize(n_cases=sum(n)) %>%

@@

```{r consumption_by_age_and_sex}
bene %>%
  left_join(total_pde, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  group_by(sex, age) %>%
  summarize(mean_n_claims=mean(n_claims)) %>%
  ungroup() %>%
  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
  geom_point()
```

@@

>>>>

## Top drugs

Which drugs are the most used?

```{r top_drugs}
top_abx = pde %>%
  count(antibiotic) %>%
  arrange(desc(n)) %>%
  head(20)

kable(top_abx)
```

@@

```{r drugs_by_sex}
usage_by_sex = function(drug) {
  bene %>%
    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id), by='bene_id') %>%
    replace_na(list(n=0)) %>%
    group_by(sex) %>%
    summarize(n_bene=n(),
              n_claims=sum(n),
              mean_n_claims=mean(n),
              sd_n_claims=sd(n)) %>%
    mutate(se_n_claims=sd_n_claims/sqrt(n_bene),
           drug=drug)
}

ubs = lapply(top_abx$antibiotic, usage_by_sex) %>% bind_rows
```

@@

```{r drug_by_sex_plot}
ubs %>%
  mutate(highlight=antibiotic %in% c('azithromycin', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'fosfomycin', 'nitrofurantoin')) %>%
  ggplot(aes(x=male, y=female, label=antibiotic, color=highlight)) +
  scale_color_manual(values=c('black', 'red')) +
  geom_abline() +
  geom_abline(slope=1.28, col='gray') +
  geom_text() +
  geom_point() +
  xlab('claims per 1,000 male beneficiaries') +
  ylab('claims per 1,000 female beneficiaries') +
  theme_minimal() +
  coord_fixed()
```

```{r drug_by_sex_plot2}
total_plot_f = function(df, highlight=character(0)) {
  df %>%
    #mutate_at(vars(total, se), mpy) %>%
    mutate(ci=se*1.96,
           ci_low=total-ci,
           ci_high=total+ci) %>%
    select(drug, sex, total, ci_low, ci_high) %>%
    gather('key', 'value', total, ci_low, ci_high) %>%
    mutate(key=case_when(.$sex=='male' & .$key=='total' ~ 'x',
                         .$sex=='male' & .$key=='ci_low' ~ 'xmin',
                         .$sex=='male' & .$key=='ci_high' ~ 'xmax',
                         .$sex=='female' & .$key=='total' ~ 'y',
                         .$sex=='female' & .$key=='ci_low' ~ 'ymin',
                         .$sex=='female' & .$key=='ci_high' ~ 'ymax')) %>%
    select(-sex) %>%
    spread(key, value) %>%
    mutate(highlight=drug %in% highlight) %>%
    ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=drug, color=highlight)) +
    geom_abline(col='gray') +
    geom_point() + geom_errorbar() + geom_errorbarh() +
    scale_color_manual(values=c('black', 'red')) +
    geom_text() +
    #geom_text(nudge_x=1, nudge_y=0.5) +
    #scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) +
    #scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 14)) +
    xlab('No. rx. for men (millions)') + ylab('No. rx. for women (millions)') +
    coord_fixed() +
    theme_minimal()
}

ubs %>%
  rename(total=mean_n_claims, se=se_n_claims) %>%
  total_plot_f()
```

@@

```{r, eval=FALSE}
# group_by_all_ = function(df, group_name) {
#   bind_rows(df, mutate_(df, .dots=setNames(list(~'all'), group_name))) %>%
#     group_by_(group_name)
# }
# 
# epi %>%
#   group_by_all_('sex') %>%
#   summarize(n_cases=sum(n)) %>%

@@

```{r consumption_by_age_and_sex, eval=FALSE}
bene %>%
  left_join(total_pde, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  group_by(sex, age) %>%
  summarize(mean_n_claims=mean(n_claims)) %>%
  ungroup() %>%
  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
  geom_point()
```

@@

## Margins of azithromycin consumption

Maybe women take more azithromycin because usage is more concentrated in women than in men. Then we could explain the increased consumption by a small number of people taking more drug. (Yonatan points to Lady Windmere syndrome.) In fact, consumption is more even among women. As always, this is driven by the extensive fraction: more women took at least one dose:

```{r, eval=FALSE}
pde %>%
  filter(drug=='azithromycin') %>%
  count(bene_id) %>% rename(n=n_azi) %>%
  right_join(bene, by='bene_id') %>% replace_na(list(n_azi=0)) %>%
  count(sex, n_azi) %>%
  group_by(sex) %>%
  mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
         cum_cons=cumsum(n_azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
  ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +
  geom_point() +
  geom_line() +
  xlab('Fraction of consumers') + ylab("Fraction of that sex's azithromycin consumed")
```
